---
title: "Data Reformating"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)

# to call stats filter- stats::filter()
```


# Reformat Catch Data


* remove "all" column
* create species column
    - move from wide to long
* general QA

[Mike Byerly. Alaska commercial salmon catches by management region (1886- 1997). Gulf of Alaska Data Portal. df35b.304.2.](https://knb.ecoinformatics.org/view/df35b.304.2)


```{r}

catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)
head(catch_original)

```

Remove the `all` and `notesRegCode` columns using `select`
Ctrl + shift +m: pipe operator shortcut %>%

```{r}
catch_long <- catch_original %>% 
            select(-All, -notesRegCode) %>%   # select columns we want
            gather(key = "species", value = "catch", -Region, -Year)  # move from wide to long

head(catch_long)

```


```{r}
catch_wide <- catch_long %>% 
    spread(key = Region, value = catch)

head(catch_wide)
View(catch_wide)

```


## Clean up our data

* rename catch to catch_thousands
* change catch column to numeric
* create a new catch column in units num. of fish

```{r}

catch_clean <- catch_long %>%
    rename(catch_thousands = catch) %>% 
    mutate(catch_thousands = ifelse(catch_thousands == "I", 1, catch_thousands)) %>% 
    mutate(catch_thousands = as.numeric(catch_thousands)) %>% 
    mutate(catch = catch_thousands * 1000) %>% 
    select(-catch_thousands)

head(catch_clean)

```


```{r, eval=FALSE, echo= FALSE}
# Look for where as.numeric failed
test <- as.numeric(catch_long$catch)
i<- which(is.na(test)==TRUE)
catch_long[i,]

```

## Split - Apply - Combine

* calculate mean catch by species

```{r}
species_mean <- catch_clean %>%
        group_by(species) %>%   
        summarise(catch_mean = mean(catch),
                  num_obs = n())

head(species_mean)
```

```{r}
year_mean <- catch_clean %>%
        group_by(Year) %>%   
        summarise(catch_mean = mean(catch)) %>% 
        arrange(-catch_mean) %>% 
        filter(Year >= 1990)

year_mean
```



# Join the Region Definitions


```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"), stringsAsFactors = FALSE) %>% 
    select(code, mgmtArea)

head(region_defs)
```


```{r}
catch_joined <- left_join(catch_clean, region_defs, by = c("Region" = "code"))

head(catch_joined)

```


# Misc. Functions

Separate and unite

```{r}
dates_df <- data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)

dates_split <- dates_df %>% 
            separate(date, c("month", "day", "year"), "/", remove=F)

head(dates_split)

```


```{r}
dates_unite <- dates_split %>% 
            unite(col = date_iso, year, month, day, sep = "-")
head(dates_unite)

```


```{r}
write.csv(catch_clean, "catch_clean.csv", row.names = F)
```

