---
title: "Data Visualization"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)

# to call stats filter- stats::filter()
```

# Reformat the data for plotting

Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1S46Q6M.


```{r}
esc <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", method = "libcurl"), stringsAsFactors = F)
```

* calculate annual escapment by species and region
* filter for just pacific salmon (Sockeye, Chinook, Coho, Pink, Chum)

annual_esc

```{r}
esc_split <- esc %>% 
            separate(sampleDate, c("year", "month", "day"), "-", remove=F)

annual_esc <- esc_split %>%
        mutate(year = as.numeric(year)) %>% 
        group_by(SASAP.Region, Species, year) %>%   
        summarise(count = sum(DailyCount)) %>% 
        filter(Species %in% c("Sockeye", "Chinook", "Coho", "Pink", "Chum"))

```



# Generate some static plots

```{r}
ggplot(annual_esc, mapping = aes(x = Species, y=count)) + geom_col(fill="blue")

ggplot(annual_esc, mapping = aes(x = Species, y=count, fill=SASAP.Region)) + geom_col()

```

```{r}
my_theme <- theme_test() + 
    theme(legend.position = "bottom", legend.title =  element_blank(),
          plot.title = element_text(hjust = 0.5))
```


```{r}
ggplot(filter(annual_esc, SASAP.Region == "Kodiak"), aes(x = year, y = count, color = Species)) + 
    geom_line() + 
    geom_point() + 
    ylab("Escapement (num of fish)") + 
    xlab("Year") + 
    ggtitle("Kodiak Salmon Escapement")
    my_theme


```

```{r, fig.width= 7, fig.height=8}
p <- ggplot(annual_esc, aes(x = year, y = count, color = Species))+
    geom_line()+
    geom_point()+
    scale_y_continuous(label=comma)+
    facet_wrap(~SASAP.Region, scales = "free_y", ncol=2)+
    my_theme

p
ggsave("figures/regional_escapement.png",
       plot=p,
       width=7,
       height=8,
       units="in")

```



# Generate an interactive map